Regression Cube Analysis of Cohort Study Data

### 2015-03-29 draft.tex
revised related work

\subsection{Abstracting Regression Results using $R^2$}
\begin{figure}[htb]
 \centering
 \includegraphics[width=3.5in]{figures/cube}
 \caption{
 (left) Overview visualization using a mosaic plot for the formula $Z \sim X + Y$, where $Z$ assumes feature $Age$.
 %%
 The $R^2$ values extracted from the regression formulas depict the quality of fit and are mapped to color saturation.
 %%
 A saturated color shows a strong correlation.
 %%
 (right) Since $Z$ assumes all features $n$ as given by the formula, it yields $n$ mosaic plot visualizations.
 %%
 These represent the slices in our cube.
 %%
 The $R^2$ values of each slice voxel is mapped on opacity in the 3D-view later on, reducing the occlusion of other values.
 }
  \label{fig:Cube}
\end{figure}
%%
The goal of an overview visualization is to provide a comprehensive view on the data (raw or using descriptive metrics \cite{Bertini}), which is easy to understand.
%% TODO: Cite another paper here!
As described in our previous work \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded using color on a mosaic plot.
%%
Regression models are more complex, having many associated describing metrics.
%%
For the \emph{Regression Cube} analysis we are interested in the quality of fit of the resulting model, which allows to infer about the predictive quality of the independent features included in the model.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the $R^2$ value is the metric allowing for this kind of assessment.
%%
\paragraph{2D (Slice) View.}
%%
Since $R^2$ is scaled between [0,1], it allows for comparison between regression models.
%%
We can apply the same mosaic plot mapping by translating the $R^2$ values to color saturation (Fig.~\ref{fig:Cube} a).
%%
This describes a 2D regression square for dynamic variables $X$ and $Y$ (e.g. $Age \sim X + Y$).
%%

@@

\textbf{Extracted Hypotheses have to investigated further.}
%%
Agreeing with TI's feedback, each finding and hypotheses has to be confirmed using dedicated statistical analysis.
%%
This incorporated investigating all potential confounding features as well as looking for outlier subjects.
%%
TI commented on the possibility of adding even more regression types to model different correlation types to customize the cube even more.
%%
%TI hinted the neccissity of different regression types, which also model non-linear relationships, but do not comprise of comparable quality of fit measures, such as the $R^2$.
%%
%\com{Tills Feedback - Regression analysis also incorporates assumptions about the data, }
%%

@@

% \com{Rewrite, focus on precise definition on what was done and why it is awesome.}
% We presented a technique for knowledge discovery in cohort study data sets with user-defined target features.
% %%
% \emph{Regression Cubes} are based on eponymous regression models, which allow to model domain knowledge about the data (e.g. influencing features on a disease) using formulas.
% %%
% Alternatively, \emph{Regression Cubes} can be calculated using default formulas to assess the explanatory power towards each feature in the data set.
% %%
% These two approaches allow both for a \emph{hypothesis-free} and \emph{hypothesis-based} explorative analysis.
% %%
% Correlation-based feature selection reduces the amount of calculations using the target feature by focusing on the important regression models.
% %%
% The prototype described in this work was developed using state of the art web technologies, free services and open source libraries.
% %%
% All code associated to this project is also open source.
% %%
% Two case studies revealed that the presented approach enables to reproduce knowledge extracted using decision-tree-based data mining methods on a hepatic steatosis data set, as well as producing new hypotheses by deriving insight into influencing factors on breast fat tissue in a explorative analysis.
% %%
%
% As future work, we want to introduce more regression models to the data set, which model different correlation types.
% %%
% We also want to extend the cube to time-dependent data by expanding the difference-cube approach.
% %%
%
% The method was very well received with our project partners, allowing them for the first time to retrieve an overview visualization of features, which does not highlights their pairwise correlations, but rather their explanatory power towards a target feature.
% %%
% The observed relationship are now subject of detailed statistical analyses.
% %%
% Our provided system not restricted to the epidemiological domain, using the provided instance, everybody can upload their own data or even set up their own \emph{Regression Cube} server cluster.
% %%
% We believe in the power of opening up knowledge discovery to allow a heterogenous group of domain experts to derive insight into their data and support the notion of open science.
%%
%%\paragraph{Future Work.}
%%\com{Implementing of regression analysis on the graphics card.}
%%
% \begin{small}
	%\acknowledgments{Omitted due to blind review}
\acknowledgments{SHIP is part of the Community Medicine Research net of the University of Greifswald, Germany, which is funded by the Federal Ministry of Education and Research (grant no. 03ZIK012), the Ministry of Cultural Affairs as well as the Social Ministry of the Federal State of Mecklenburg-West Pomerania. Whole-body MR imaging was supported by a joint grant from Siemens Healthcare, Erlangen, Germany and the Federal State of Mecklenburg-Vorpommern. The University of Greifswald is a member of the ‘Centre of Knowledge Interchange’ program of the Siemens AG. This work was supported by the DFG Priority Program 1335: Scalable Visual Analytics.}
% \end{small}
\clearpage
\newpage
\bibliographystyle{abbrv}
%%use following if all content of bibtex file should be shown
%\nocite{*}
\bibliography{bibliography}
\end{document}

>>>>

\subsection{Abstracting Regression Results using $R^2$}
\begin{figure}[htb]
 \centering
 \includegraphics[width=3.5in]{figures/cube}
 \caption{
 (left) Overview visualization using a mosaic plot for the formula $Z \sim X + Y$, where $Z$ assumes feature $Age$.
 %%
 The $R^2$ values extracted from the regression formulas depict the quality-of-fit and are mapped to color saturation.
 %%
 A saturated color shows a strong correlation.
 %%
 (right) Since $Z$ assumes all features $n$ as given by the formula, it yields $n$ mosaic plot visualizations.
 %%
 These represent the slices in our cube.
 %%
 The $R^2$ values of each slice voxel is mapped on opacity in the 3D-view later on, reducing the occlusion of other values.
 }
  \label{fig:Cube}
\end{figure}
%%
The goal of an overview visualization is to provide a comprehensive view on the data (raw or using descriptive metrics \cite{Bertini}), which is easy to understand.
%% TODO: Cite another paper here!
As described in our previous work \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded using color on a mosaic plot.
%%
Regression models are more complex, having many associated describing metrics.
%%
For the \emph{Regression Cube} analysis we are interested in the quality-of-fit of the resulting model, which allows to infer about the predictive quality of the independent features included in the model.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the $R^2$ value is the metric allowing for this kind of assessment.
%%
\paragraph{2D (Slice) View.}
%%
Since $R^2$ is scaled between [0,1], it allows for comparison between regression models.
%%
We can apply the same mosaic plot mapping by translating the $R^2$ values to color saturation (Fig.~\ref{fig:Cube} a).
%%
This describes a 2D regression square for dynamic variables $X$ and $Y$ (e.g. $Age \sim X + Y$).
%%

@@

\textbf{Extracted Hypotheses have to investigated further.}
%%
Agreeing with TI's feedback, each finding and hypotheses has to be confirmed using dedicated statistical analysis.
%%
This incorporated investigating all potential confounding features as well as looking for outlier subjects.
%%
TI commented on the possibility of adding even more regression types to model different correlation types to customize the cube even more.
%%
%TI hinted the neccissity of different regression types, which also model non-linear relationships, but do not comprise of comparable quality-of-fit measures, such as the $R^2$.
%%
%\com{Tills Feedback - Regression analysis also incorporates assumptions about the data, }
%%

@@

% \com{Rewrite, focus on precise definition on what was done and why it is awesome.}
% We presented a technique for knowledge discovery in cohort study data sets with user-defined target features.
% %%
% \emph{Regression Cubes} are based on eponymous regression models, which allow to model domain knowledge about the data (e.g. influencing features on a disease) using formulas.
% %%
% Alternatively, \emph{Regression Cubes} can be calculated using default formulas to assess the explanatory power towards each feature in the data set.
% %%
% These two approaches allow both for a \emph{hypothesis-free} and \emph{hypothesis-based} explorative analysis.
% %%
% Correlation-based feature selection reduces the amount of calculations using the target feature by focusing on the important regression models.
% %%
% The prototype described in this work was developed using state of the art web technologies, free services and open source libraries.
% %%
% All code associated to this project is also open source.
% %%
% Two case studies revealed that the presented approach enables to reproduce knowledge extracted using decision-tree-based data mining methods on a hepatic steatosis data set, as well as producing new hypotheses by deriving insight into influencing factors on breast fat tissue in a explorative analysis.
% %%
%
% As future work, we want to introduce more regression models to the data set, which model different correlation types.
% %%
% We also want to extend the cube to time-dependent data by expanding the difference-cube approach.
% %%
%
% The method was very well received with our project partners, allowing them for the first time to retrieve an overview visualization of features, which does not highlights their pairwise correlations, but rather their explanatory power towards a target feature.
% %%
% The observed relationship are now subject of detailed statistical analyses.
% %%
% Our provided system not restricted to the epidemiological domain, using the provided instance, everybody can upload their own data or even set up their own \emph{Regression Cube} server cluster.
% %%
% We believe in the power of opening up knowledge discovery to allow a heterogenous group of domain experts to derive insight into their data and support the notion of open science.
%%
%%\paragraph{Future Work.}
%%\com{Implementing of regression analysis on the graphics card.}
%%
% \begin{small}
	%\acknowledgments{Omitted due to blind review}
\acknowledgments{SHIP is part of the Community Medicine Research net of the University of Greifswald, Germany, which is funded by the Federal Ministry of Education and Research (grant no. 03ZIK012), the Ministry of Cultural Affairs as well as the Social Ministry of the Federal State of Mecklenburg-West Pomerania. Whole-body MR imaging was supported by a joint grant from Siemens Healthcare, Erlangen, Germany and the Federal State of Mecklenburg-Vorpommern. The University of Greifswald is a member of the âCentre of Knowledge Interchangeâ program of the Siemens AG. This work was supported by the DFG Priority Program 1335: Scalable Visual Analytics.}
% \end{small}
\clearpage
\newpage
\bibliographystyle{abbrv}
%%use following if all content of bibtex file should be shown
%\nocite{*}
\bibliography{bibliography}
\end{document}
